---
title: "Results Vanished Journals"
output: github_document
---

```{r, echo = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  fig.width = 6,
  fig.asp = 0.618,
  out.width = "70%",
  fig.align = "center",
  dpi = 300
)
```

## Load data

Data was manually compiled and curated using a Google spreadsheet.

<https://docs.google.com/spreadsheets/d/1M8WdNFpT8Xd-Bj7Eo_4Qcj-jBOdKQhRNxOHj1ckC0n0/edit#gid=0>

```{r}
library(tidyverse)
library(googledrive)
library(cowplot)
googledrive::drive_download(
  googledrive::as_id("https://docs.google.com/spreadsheets/d/1M8WdNFpT8Xd-Bj7Eo_4Qcj-jBOdKQhRNxOHj1ckC0n0/edit#gid=0"), 
  path = "data/vanished_journals.xlsx",
  overwrite = TRUE
  )
jn_raw <- readxl::read_xlsx("data/vanished_journals.xlsx")
#' date transformation
my_jn <- jn_raw %>%
  mutate_at(vars(contains("Year")), as.character) %>%
  mutate_at(vars(contains("Year")), function(x) gsub(".0", "", x, fixed = TRUE)) %>%
  mutate_at(vars(contains("Year")), function(x) lubridate::parse_date_time(x, c('y')))
```

### How many journals disappeared?

```{r}
my_jn %>%
  distinct(`Journal Name`) %>%
  nrow()
```

### When did they vanished from the Web?


```{r}
my_jn %>% 
filter(!is.na(`Last Year Online`)) %>%
  ggplot(aes(as.Date(`Last Year Online`, origin = "%Y-%m-%d"))) +
  geom_histogram(stat = "count") +
  scale_x_date(date_labels = "%Y", 
               limits = c(as.Date("2000-01-01"), as.Date("2019-12-31")),
               date_breaks = "3 year"
               ) +
  labs(x = "", y = "Journals") +
  theme_minimal_hgrid(12)
```

### For how long were the journals actively publishing?

```{r}
my_jn <- my_jn %>%
  mutate(publishing_period = lubridate::year(`End Year`) - lubridate::year(`Start Year`)) %>%
  filter(publishing_period > 1)

ggplot(my_jn, aes(x = publishing_period)) +
  geom_histogram(binwidth = 4, fill="#56B4E9", alpha=0.5, color = "#56B4E9") +
  geom_vline(aes(xintercept = mean(publishing_period, na.rm = T)),
             colour = "#E69F00", linetype ="dashed", size = .8) +
  geom_vline(aes(xintercept = median(publishing_period, na.rm = T)),
             colour = "red", linetype ="dashed", size = .8) +
  theme_minimal_hgrid(12) +
  scale_x_continuous("Years active") +
  ylab("Journals") 
```

### Life-span of ceased journals 

```{r}
my_jn_timeline <- my_jn %>%
  mutate(start_year = lubridate::year(`Start Year`),
         end_year = lubridate::year(`End Year`)) %>%
  filter(!is.na(end_year)) %>%
  filter(end_year != start_year) %>%
  ungroup() %>%
  mutate(jn = fct_reorder(as.factor(`Journal Name`), start_year)) 




ggplot(my_jn_timeline, aes(ymin = start_year, 
                  ymax = end_year, 
                  x = reorder(jn, -end_year), 
                  color = end_year)) +
  coord_flip(ylim = c(1990, 2018)) +
  geom_linerange() +
  labs(x = NULL) +
  scale_color_viridis_c(option = "C", begin = 0, end = 0.8) +
  scale_y_continuous(breaks = seq(1990, 2020, by =4)) +
  theme_minimal_vgrid(12) +
  theme(axis.text.y = element_blank(),
            axis.ticks.y = element_blank()) +
  theme(legend.position="none")
```

#### by Discipline

Summary table

```{r, results='asis'}
my_jn %>%
  filter(!is.na(`Discipline (where more info available)`), `Discipline (where more info available)` != "?") %>%
 rename(subject = `Discipline (where more info available)`) %>% 
  group_by(subject) %>%
  summarise(n_journals = n_distinct(`Journal Name`),
            mean = mean(publishing_period),
            median = median(publishing_period),
            q25 = quantile(publishing_period, probs = .25),
            q75 = quantile(publishing_period, probs = .75)) %>%
  knitr::kable()
```



Classic boxplot

```{r}
my_jn %>%
  filter(!is.na(`Discipline (where more info available)`), `Discipline (where more info available)` != "?") %>%
  rename(subject = `Discipline (where more info available)`) %>%
  ggplot(aes(subject, publishing_period, color = subject)) +
  geom_boxplot(varwidth = TRUE) +
  geom_jitter(width = 0.2, alpha = 0.5) +
  theme_minimal_hgrid(12) +
  labs(x = NULL, y = "Years active") +
  coord_cartesian(ylim = c(0, 20)) +
  scale_colour_brewer(type = "qual", palette = "Set1") +
  theme(legend.position="none")
```


#### by country

Journals originate from `r length(unique(my_jn$Country))` distinct countries.

```{r, results='asis'}
my_jn %>%
  count(Country, sort = TRUE) %>%
  mutate(prop = n / sum(n) * 100) %>%
  knitr::kable()
```

per Word Bank region

```{r, results='asis'}
library(countrycode)
my_jn %>%
  mutate(region = countrycode::countrycode(Country, "country.name", "region")) %>%
  count(region, sort = TRUE) %>%
  mutate(prop = n / sum(n) * 100) %>%
  knitr::kable()
```

Suggestion: make a world map similar to

<https://github.com/subugoe/oa2020cadata/blob/master/analysis/paper.md#global-map-of-journal-publishing>

where circle size represents the number of journals, an pies represent the subject

```{r}
map_data <-my_jn %>%
  filter(!is.na(`Discipline (where more info available)`), `Discipline (where more info available)` != "?") %>%
  rename(subject = `Discipline (where more info available)`) %>%
  mutate(region = countrycode::countrycode(Country, "country.name", "region")) %>%
  count(region, subject, sort = TRUE) 
# manual lat long mapping regions
# map_data %>%
#   distinct(region) %>%
#   write_csv("region_lat_long.csv")
latlong <- readr::read_csv("region_lat_long.csv")
tmp <- inner_join(map_data, latlong, by = "region") %>%
  spread(subject, n) %>%
  mutate_at(c(4:7), ~replace(., is.na(.), 0)) %>%
  mutate(all_n = Health + LS + PSM + SSH) 
world <- map_data("world")  %>%
  # without Antartica
  filter(region != "Antarctica")
p <- 
  ggplot(world, aes(long, lat)) +
  geom_map(
    map = world,
    aes(map_id = region),
    color = "white",
    size = 0.1,
    fill  = "#b3b3b3a0"
  ) +
  coord_sf() 
p +
  scatterpie::geom_scatterpie(
    aes(long, lat, r = sqrt(all_n) * 3),
    data = tmp,
    cols = c("Health", "LS", "PSM", "SSH"),
    alpha = .7,
    color = NA
  )  +
 # scale_fill_brewer("Journal Subject", palette = "Set1") +
  scale_fill_manual("Journal Subject", values = c("#FAAB18", "#1380A1","#990000", "#588300")) +
  labs(x = NULL, y = NULL) +
  coord_equal() +
  theme_minimal() +
  theme(plot.margin = margin(10, 30, 10, 30)) +
  theme(panel.grid.minor = element_blank()) +
  theme(axis.ticks = element_blank()) +
  theme(panel.grid.major = element_blank()) +
  theme(panel.border = element_blank()) +
  theme(axis.ticks = element_blank(),
        axis.text = element_blank()) +
  theme(legend.position = "top",
        legend.justification = "center") +
  theme(legend.text = element_text(size = 7.5))
```

#### by evidence source

```{r, results='asis'}
my_jn %>%
  count(Source, sort = TRUE) %>%
  mutate(prop = n / sum(n) * 100) %>%
  knitr::kable()
```

### Archives

The Keepers by subject

```{r, results='asis'}
my_jn %>%
   filter(!is.na(`Discipline (where more info available)`), `Discipline (where more info available)` != "?") %>%
  rename(subject = `Discipline (where more info available)`) %>%
  group_by(subject, `The Keepers (archived)`) %>%
  summarise(
    n_journals = n_distinct(`Journal Name`)
    ) %>%
  knitr::kable()
```
